---
title: "p8105_hw2_xl3141"
author: "Xinyuan Liu"
date: "2021/10/1"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
mr_trash_wheel_df <-
  read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "A2:N408", sheet = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(across(sports_balls, round, 0))
```

```{r, median, echo=TRUE, eval=FALSE}
wheel_median <- 
   mr_trash_wheel_df %>% 
   select(year, sports_balls) %>% 
   filter(year == "2017") 
median(pull(wheel_median, sports_balls))
```

In the Mr. trash wheel dataset, we have in total `r nrow(mr_trash_wheel_df)` rows and `r ncol(mr_trash_wheel_df)` columns. The variables that we are interested includes `r names(mr_trash_wheel_df)`. The median number of sports ball in a dumpster in 2017 is 8.

```{r }
precip_2019_df <- 
  read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "A2:B14", sheet = 4) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2019")

precip_2018_df <- 
  read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "A2:B15", sheet = 5) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2018")
  
precip_1819_df <- 
  bind_rows(precip_2018_df,precip_2019_df) %>% 
  mutate(month = month.name[month])

```

We have `r nrow(precip_1819_df)` months' precipitation data for 2018 and 2019 in total. The total precipitation in 2018 is `r sum(pull(precip_2018_df, total))`, and the total precipitation in 2019 is `r sum(pull(precip_2019_df, total))`.