---
title: "p8105_hw2_xl3141"
author: "Xinyuan Liu"
date: "2021/10/1"
output: github_document
---
## Problem 1

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
mr_trash_wheel_df <-
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", range = "A2:N535", sheet = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% ##omit rows that do not include dumpster-specific data
  mutate(across(sports_balls, round, 0)) ##round the number of sports balls to the nearest integer
```

```{r, median, echo=TRUE, eval=FALSE}
wheel_median <- 
   mr_trash_wheel_df %>% 
   select(year, sports_balls) %>% 
   filter(year == "2017") 
median(pull(wheel_median, sports_balls)) ## calculate the median number of sports balls in a dumpster in 2017
```

In the Mr. trash wheel dataset, we have in total `r nrow(mr_trash_wheel_df)` rows and `r ncol(mr_trash_wheel_df)` columns. The variables that we are interested includes _`r names(mr_trash_wheel_df)`_. The median number of sports ball in a dumpster in 2017 is 8.

```{r message=FALSE, warning=FALSE}
precip_2019_df <- 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", range = "A2:B15", sheet = 6) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% ## drop na rows
  mutate(year = "2019") ##add variable "year"

precip_2018_df <- 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", range = "A2:B15", sheet = 7) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% ## drop na rows
  mutate(year = "2018") ##add variable "year"
  
precip_1819_df <- 
  bind_rows(precip_2018_df,precip_2019_df) %>% 
  mutate(month = month.name[month]) ##combine precipitation datasets and convert month to a character variable 

```

We have `r nrow(precip_1819_df)` months' precipitation data for 2018 and 2019 in total. The total precipitation in 2018 is `r sum(pull(precip_2018_df, total))`, and the total precipitation in 2019 is `r sum(pull(precip_2019_df, total))`.


## Problem 2

```{r message=FALSE, warning=FALSE}
pol_month_df <- 
  read_csv("fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% ##separate date into month, day, and year
  mutate(month = month.name[as.numeric(month)]) %>% ## change numeric month into character
  mutate(president = prez_gop + prez_dem) %>% ## merge columns "prez_gop" and "prez_dem"
  select(-prez_gop, -prez_dem, -day)##remove column"prez_gop", "prez_dem", "day"
  
```

"pol" dataset describes the number of national politicians who are democratic or republican at any given time. It has `r nrow(pol_month_df)` rows and `r ncol(pol_month_df)` columns. The variable includes _`r names(pol_month_df)`_. It includes the data from 1974 to 2015.

```{r message=FALSE, warning=FALSE}
snp_df <- 
  read_csv("fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names()%>% 
  separate(date, into = c("month", "day", "year")) %>%  ##separate date into month, day, and year
  select(-day) %>% ##remove "day" column
  relocate("year", "month") %>% 
  mutate(
    year = as.numeric(year), 
    year = ifelse(year > 40, year + 1900, year + 2000),
    month = month.name[as.numeric(month)],
    year = as.character(year)##make both month and year into character variable in order to merge the dataframe
  )
range(pull(snp_df, year))
  
unemploy_df <-
  read_csv("fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
    ) %>%  ## make the month into one column
  relocate("year", "month") %>% 
  mutate(
    month = str_to_title(month), 
    month = match(month, month.abb),
    month = month.name[as.numeric(month)],
    year = as.character(year) ##make both month and year into character variable in order to merge the dataframe
    )
```

```{r message=FALSE, warning=FALSE}
new_df_1 <- left_join(pol_month_df, snp_df, by = c("year" = "year", "month" = "month")) ## join "snp_df" into "pol_month_df"
result_df <- left_join(new_df_1, unemploy_df, by = c("year" = "year", "month" = "month")) ## join the "unemploy_df" into "new_df_1"
range(pull(result_df, year))
```

The result dataset has `r nrow(result_df)` rows and `r ncol(result_df)` columns. The variables includes _`r names(result_df)`_. It includes the data from 1947 to 2015.


## Problem 3
```{r message=FALSE, warning=FALSE}
pop_baby_name_df <-
  read_csv("Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    childs_first_name = str_to_title(childs_first_name), ## capitalize the first letter
    ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC")  ##make the name of same variable identical
    ) %>% 
  arrange(gender, ethnicity) %>% ## arrange according to gender and ethnicity
  distinct() ## remove duplicated rows
  
olivia_fem_pop_df <- 
  filter(pop_baby_name_df, childs_first_name == "Olivia" ) %>% 
  select(-gender, -childs_first_name, -count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>% 
  knitr::kable() ## produce the table for the popularity of "Olivia" in female for all ethnicity from 2011 to 2016
olivia_fem_pop_df

male_pop_df <- 
  filter(pop_baby_name_df, gender == "MALE", rank <= 5) %>% 
  select(-gender, -count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>% 
  knitr::kable() ## produce the table for the top 5 names in male for all ethnicity from 2011 to 2016
male_pop_df

name_vs_pop_df <-
  pop_baby_name_df %>% 
  filter(year_of_birth == "2016", ethnicity == "WHITE NON HISPANIC", gender == "MALE") ##dataset with names of male white non-hispanic children born in 2016

ggp_rank_num <- 
  ggplot(name_vs_pop_df, aes(x = rank, y = count)) + 
  geom_point() + xlab("rank in popularity") + 
  ylab("number of children") + 
  ggtitle("popularity of names for male white non-hispanic children born in 2016") ##scatterplot for number of children with a name against rank in popularity of that name
ggsave("scatter_plot.pdf")
```

